#TensorDecomp Based Classification on BNU1 dataset

In this experiment, we carry out tensor decomposition on one dataset and obtain low dimensional representation of 46 subjects. The goal is to assess if TensorDecomp results in any loss (if not improvement) in discriminability wrt labels (gender in this case).


## Tensor decomp

Tensor decomp first, without using any info about labels.

```{r}
require("TensorEmbedding")
require("base")
require("rPython")
require("plyr")

hash<- function(x){
  python.exec( "def h(a): return abs(hash(a))%100000" )
  python.call( "h", x)
}


setwd("~/git/semipar-cci/")

#load processed data
load("./Data/processed/KKI2009.RDa")

d<-dataList

SEX<- numeric()

m<- length(d)
n<- nrow(d[[1]]$A)
id<- numeric()


A<- numeric()
for(j in 1:m){
    A<- cbind(A, c(d[[j]]$A))
    SEX<- c(SEX, d[[j]]$SEX)
        id<- c(id, hash(paste(d[[j]]$id)))
}

# id<- unlist(lapply(d, function(x){x$id}))

k<- 30

#tensor decomp:
# fit <- TensorEmbedding::symm_group_tensor_decomp(A, id, n, m, k, 1000, 1E-5, 1E-5)
# save(fit,file="usecase/KKI2009.Rda")

#let's load the serialized one for now
load(file="usecase/KKI2009.Rda")
```


Plot the eigenvalues, colored by labels
```{r}

SEXwUID<- tapply(SEX, id, function(x){x[1]})
uid0<- unique(id)
sUID<- order(uid0)
UID<- uid0[sUID]
m<- length(uid0)

vectorsSortedByUID<- fit$diagC[sUID,]


plot(c(1,k),range(vectorsSortedByUID),type="n")
for(i in 1:m){
  lines(vectorsSortedByUID[i,],col=SEX[i])
}
```

## Classification


###K-nearest neighbor


KNN function:

```{R}

knn<- function(i,dist,k=10,cl){
  o<- order(dist[i,])
  ct<-count(cl[o>1 & o<=(k+1)])
  majority_vote<- ct$x[ct$freq == max(ct$freq)] 
  if(length(majority_vote)>1)
    sample(x = majority_vote,size = 1)
  else
    majority_vote
}

knn_mc_rate <-function(dist,k){
  sum(SEXwUID != sapply(c(1:m), function(x){knn(x,dist,k,SEXwUID)}))/m
}


```




1. On TensorDecomp-produced vector

```{R}

dist<- matrix(0,m,m)
for(i in 1:m){
  for(j in 1:(i)){
    dist[i,j]<- sqrt(sum((vectorsSortedByUID[i,]-vectorsSortedByUID[j,])^2))
    dist[j,i]<- dist[i,j]
  }
}


knn_rate_vs_k<- sapply(c(1:30), function(k)knn_mc_rate(dist,k))

plot(c(1:30),knn_rate_vs_k, type="l",xlab="k",ylab="C14 Error Rate", col="red")

min(knn_rate_vs_k)

```



2. On full adjacency matrices

```{R}

countByID<- tapply(id, id, function(x){length(x)})
subOnePerIdx<- cumsum(countByID)


matA<- t(A[,order(id)][,subOnePerIdx])

dist<- matrix(0,m,m)
for(i in 1:m){
  for(j in 1:(i)){
    dist[i,j]<- sqrt(sum((matA[i,]-matA[j,])^2))
    dist[j,i]<- dist[i,j]
  }
}


knn_rate_vs_k<- sapply(c(1:(30)), function(k)knn_mc_rate(dist,k))

plot(c(1:(30)),knn_rate_vs_k, type="l",xlab="k",ylab="C14 Error Rate", col="red")

min(knn_rate_vs_k)

```