#TensorDecomp Based Classification on 4 datasets



## Tensor decomp

Tensor decomp first, without using any info about labels.

```{r}
require("TensorEmbedding")
require("base")
require("rPython")
require("plyr")
  require("GGally")
 require("ggplot2")

hash<- function(x){
  python.exec( "def h(a): return abs(hash(a))%100000" )
  python.call( "h", x)
}

setwd("~/git/semipar-cci/")

loadData<- function(dataset){
  load(file=paste("Data/processed/",dataset,".RDa",sep=""))
  dataList
}


BNU1<-loadData("BNU1")
BNU3<-loadData("BNU3")
KKI2009<-loadData("KKI2009")
MRN114<-loadData("MRN114")


hash("BNU1")

data4sets<- list(BNU1,BNU3, KKI2009, MRN114)

A<- numeric()
id<- numeric()
batchID<- numeric()
SEX<- numeric()

for(i in 1:length(data4sets)){
  d<- data4sets[[i]]
  
  m<- length(d)
  n<- nrow(d[[1]]$A)
  
  for(j in 1:m){
    A<- cbind(A, c(d[[j]]$A))
    id<- c(id, hash(paste(i,d[[j]]$id)))
    batchID<- c(batchID,i) 
    SEX<- c(SEX, d[[j]]$SEX)
  }
}


m<- ncol(A)
n<- sqrt(nrow(A))

k<- 30
# fit <- TensorEmbedding::symm_group_tensor_decomp(A, id, n, m, k, 500, 1E-5, 1E-5)

# save(fit,file="usecase/4datasetFit.RDa")
# save(id,file="usecase/4datasetFitID.RDa")


#let's load the serialized one for now
load(file="usecase/4datasetFit.RDa")
load(file="usecase/4datasetFitID.RDa")

```



## Batch effects by data sources

Batch means
```{r}

#batchID with Unique subject IDs(sorted)
batchIDwUID<- tapply(batchID, id, function(x){x[1]})
SEXwUID<- tapply(SEX, id, function(x){x[1]})

uid0<- unique(id)
sUID<- order(uid0)
UID<- uid0[sUID]

vectorsSortedByUID<- fit$diagC[sUID,]


plot(c(1,k),range(fit$diagC),type="n")
for(i in 1:4){
  lines(colMeans(vectorsSortedByUID[batchIDwUID==i,]),col=i,lwd=1)
}
```


Plot the eigenvalues, colored by labels
```{r}

m<- length(unique(id))

plot(c(1,k),range(vectorsSortedByUID),type="n")

for(i in 1:m){
  lines(vectorsSortedByUID[i,],col=SEXwUID[i])
}
```


Plot the eigenvalues, colored by labels
```{r}


CharSex =  as.character(SEXwUID)

dataPairs <- data.frame("ev"=vectorsSortedByUID, "SEX"=as.factor(CharSex))
ggpairs(dataPairs,columns = c(1:5), aes(colour = SEX, alpha=0.4))
```


Remove batch mean from each subject

```{r}

#batchID with Unique subject IDs(sorted)
vectorsBEremoved<- vectorsSortedByUID
for(i in 1:4){
  me<- colMeans(vectorsSortedByUID[batchIDwUID==i,])
  vectorsBEremoved[batchIDwUID==i,]<-
      t(t(vectorsBEremoved[batchIDwUID==i,])-me)
}

m<- nrow(vectorsBEremoved)

plot(c(1,k),range(vectorsBEremoved),type="n")
for(i in 1:m){
  lines(vectorsBEremoved[i,],col=SEXwUID[i])
}


```


## Classification


###K-nearest neighbor


KNN function:

```{R}

knn<- function(i,dist,k=10,cl){
  o<- order(dist[i,])
  ct<-count(cl[o>1 & o<=(k+1)])
  majority_vote<- ct$x[ct$freq == max(ct$freq)] 
  if(length(majority_vote)>1)
    sample(x = majority_vote,size = 1)
  else
    majority_vote
}

knn_mc_rate <-function(dist,k){
  sum(SEXwUID != sapply(c(1:m), function(x){knn(x,dist,k,SEXwUID)}))/m
}


```



1. On TensorDecomp-produced vector (without Batch Effects)

```{R}


dist<- matrix(0,m,m)
for(i in 1:m){
  for(j in 1:(i)){
    dist[i,j]<- sqrt(sum((vectorsBEremoved[i,]-vectorsBEremoved[j,])^2))
    dist[j,i]<- dist[i,j]
  }
}


knn_rate_vs_k<- sapply(c(1:30), function(k)knn_mc_rate(dist,k))

plot(c(1:30),knn_rate_vs_k, type="l",xlab="k",ylab="C14 Error Rate", col="red")

min(knn_rate_vs_k)

```


2. On TensorDecomp-produced vector (with Batch Effects)

```{R}


dist<- matrix(0,m,m)
for(i in 1:m){
  for(j in 1:(i)){
    dist[i,j]<- sqrt(sum((vectorsSortedByUID[i,]-vectorsSortedByUID[j,])^2))
    dist[j,i]<- dist[i,j]
  }
}


knn_rate_vs_k<- sapply(c(1:30), function(k)knn_mc_rate(dist,k))

plot(c(1:30),knn_rate_vs_k, type="l",xlab="k",ylab="C14 Error Rate", col="red")

min(knn_rate_vs_k)

```


3. On full adjacency matrices

```{R}

countByID<- tapply(batchID, id, function(x){length(x)})
subOnePerIdx<- cumsum(countByID)


matA<- t(A[,order(id)][,subOnePerIdx])

dist<- matrix(0,m,m)
for(i in 1:m){
  for(j in 1:(i)){
    dist[i,j]<- sqrt(sum((matA[i,]-matA[j,])^2))
    dist[j,i]<- dist[i,j]
  }
}


knn_rate_vs_k<- sapply(c(1:(30)), function(k)knn_mc_rate(dist,k))

plot(c(1:(30)),knn_rate_vs_k, type="l",xlab="k",ylab="C14 Error Rate", col="red")

min(knn_rate_vs_k)

```
