---
title: "Title"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---


```{r,echo=FALSE}
require('igraph')
require('ggplot2')
require('reshape')
setwd("~/git/ABadea/")

listGs<- list.files(path = "./graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("./covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

#load and sort the tensor decomp data
load("./tensorDecompBinary.Rda")
load("./listData.Rda")
ids2 <- unlist( lapply(dataList,function(x)strtrim(x$id,6)))
diagC<- tensorDecomp$C[ids2%in% common_id,]
diagC<- diagC[order(ids2[ids2%in% common_id]),]



graphList<- lapply(listGs, function(x){
  read.graph( file = paste("./graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

```


```{r,echo=FALSE}


order_by_hemi<- order(HemisphereList[[1]])


AbyHemiSphere<- (as.matrix(AdjacencyList[[1]]))[order_by_hemi,order_by_hemi]

plotHeatmap<-function(denseA, zlimMax=1){
  rownames(denseA) <- c(1:nrow(denseA))
  colnames(denseA) <- c(1:nrow(denseA))
  m<-melt(denseA)
  n<- nrow(denseA)
  p <- ggplot(m, aes(X1, X2)) +
    geom_tile(aes(fill =  value), colour = "white") + 
    scale_fill_gradient(low = "white",  high = "red",limits=c(0,zlimMax)) +
    theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.y = element_blank()) 
  
  p
}


m<- length(AdjacencyList)


n<- nrow(AdjacencyList[[1]])

computeAvgA<- function(l){
  m<- length(l)
  sumAdj<- matrix(0,n,n)
  for(i in 1:m){
   sumAdj<- sumAdj + (l[[i]]) [order_by_hemi,order_by_hemi]
  }
  as.matrix(sumAdj/m)
}

averageAdjacency<- computeAvgA(AdjacencyList)

```

#Age related degeneration


### The avg.degree vs age (weeks)

```{r}

weeks<- covariates$AGE_WEEKS

degreeForAll<-c(sapply(AdjacencyList,function(x){
  rowSums(as.matrix(x))
}))


df<- data.frame("ROI_Index_L_and_R"= rep(c(1:n),m),
                "Age"= as.factor(rep(weeks,each=n)),
                "Degree" = degreeForAll)

ggplot(data=df,
       aes(x=ROI_Index_L_and_R, y=Degree, colour=Age)) +
  geom_line()

ggplot(df, aes(x=Degree, fill=Age)) + geom_density(alpha = 0.7, bw = 5)

```

Plot total degree over age:

```{r}

sumAForAll<-c(sapply(AdjacencyList,function(x){
  sum(as.matrix(x))
}))

degreeForAll<-(lapply(AdjacencyList,function(x){
  rowSums(as.matrix(x))
}))

nnz<-sapply(degreeForAll, function(x){sum(x>0)})

df<- data.frame("Age"= weeks,
                "AvgDegree" = sumAForAll/n,
                "AvgDegreeNZ" = sumAForAll/nnz,
                "GenoType" = as.factor(covariates$GENOTYPE),
                "Gender" = as.factor(covariates$GENDER)
                )


ggplot(data=df,
       aes(x=Age, y=AvgDegree,col=GenoType)) +
  geom_line()+ggtitle("AvgDegree vs Age")

ggplot(data=df,
       aes(x=Age, y=AvgDegree,col=Gender)) +
  geom_line()+ggtitle("AvgDegrees vs Age")


ggplot(data=df,
       aes(x=Age, y=AvgDegreeNZ,col=GenoType)) +
  geom_line()+ggtitle("AvgDegree (non-zero) vs Age")

ggplot(data=df,
       aes(x=Age, y=AvgDegreeNZ,col=Gender)) +
  geom_line()+ggtitle("AvgDegrees (non-zero) vs Age")

```

### Divide age into 3 groups

We use 25% (63) and 75% (78) quantiles as cutoffs to divide mice into three group:
*young*,*middle* and *old*.
```{r}
plot(weeks)
abline(h=63,lty=2)
abline(h=78,lty=2)

quantile(weeks,c(0.25,0.75))
```

### Comparing Adjacency matrices

Obviously, *old group seems significantly more degenerate than the young and middle*, while the difference beween young and middle is more subtle.

```{r,echo=FALSE}

youngAvgA <- computeAvgA(AdjacencyList[weeks<=63])
middleageAvgA <- computeAvgA(AdjacencyList[weeks>63 & weeks<= 78  ])
oldAvgA <- computeAvgA(AdjacencyList[weeks>78])

zlim<- max(c(youngAvgA, middleageAvgA, oldAvgA))

youngAvgWholeBrain<- plotHeatmap(youngAvgA,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("Young")

middleageAvgWholeBrain<- plotHeatmap(middleageAvgA,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("Middle")

oldAvgWholeBrain<- plotHeatmap(oldAvgA,zlimMax = zlim) + geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("old") 

```


Young (n=):
```{r,echo=FALSE}
sum(weeks<=63)
youngAvgWholeBrain
```

Middle (n=):
```{r,echo=FALSE}
sum(weeks>63 & weeks<= 78)

middleageAvgWholeBrain

```

Old (n=):
```{r,echo=FALSE}
sum(weeks> 78)
oldAvgWholeBrain
```


<!-- The dichotomized view with connecting probability>0: it shows the old clearly has less connectivity than the middleage and young. -->

<!-- ```{r,echo=FALSE} -->

<!-- zlim<- 1 -->

<!-- youngAvgWholeBrainDicho<- plotHeatmap((youngAvgA>0)*1,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+ -->
<!--   ggtitle("Young") -->

<!-- middleageAvgWholeBrainDicho<- plotHeatmap( (middleageAvgA>0)*1,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+ -->
<!--   ggtitle("Middle") -->

<!-- oldAvgWholeBrainDicho<- plotHeatmap((oldAvgA>0)*1,zlimMax = zlim) + geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+ -->
<!--   ggtitle("old")  -->

<!-- youngAvgWholeBrainDicho -->
<!-- middleageAvgWholeBrainDicho -->
<!-- oldAvgWholeBrainDicho -->

<!-- ``` -->

###Degree distribution of the 3 age groups:

Middle and young mice have higher level of connectivy  (degree) than the old.

The degree in the average adjacency:

```{r,echo=FALSE}

df<- data.frame("ROI_Index"= rep(c(1:(n)),3),
                "Age"=as.factor(rep(c("Young","Middle","Old"),each=n)),
                "Degree" = c(rowSums(youngAvgA),rowSums(middleageAvgA),rowSums(oldAvgA))
)

df$Age<- ordered(df$Age, levels = c("Young","Middle","Old"))

# df<-df[df$Age!="Young",]

ggplot(df, aes(x=Degree, fill=Age)) + geom_density(alpha = 0.7, bw = 10)

```

```{r,echo=FALSE}

ageGroup <-  (weeks<=63)*1 + (weeks>63 & weeks<=78)*2+(weeks>78)*3
ageGroup[ageGroup==1] = "Young"
ageGroup[ageGroup==2] = "Middle"
ageGroup[ageGroup==3] = "Old"

df<- data.frame(
  "Age"= c("Young","Middle","Old"),
  "AvgDegree"=c(sum((youngAvgA)),sum((middleageAvgA)*1),sum((oldAvgA)*1))/n
  )

df$Age<- ordered(df$Age, levels = c("Young","Middle","Old"))


ggplot(data=df, aes(x=Age, y=AvgDegree, fill=Age)) +
    geom_bar(stat="identity")
    
```

```{r,echo=FALSE}

getAvgDegree<- function(l){
  sapply(l, function(x) mean(rowSums(as.matrix(x))))
}

getAvgDegree(AdjacencyList[weeks<=63])

youngAvgAList <- getAvgDegree(AdjacencyList[weeks<=63])
middleageAvgAList <- getAvgDegree(AdjacencyList[weeks>63 & weeks<= 78  ])
oldAvgAList <- getAvgDegree(AdjacencyList[weeks>78])

n1<- length(youngAvgAList)
n2<- length(middleageAvgAList)
n3<- length(oldAvgAList)


df<- data.frame(
  "x" = runif(n1+n2+n3,-1,1),
  "Age"= as.factor(c(rep("Young",n1),rep("Middle",n2),rep("Old",n3))),
  "AvgDegree"=c(youngAvgAList,middleageAvgAList,oldAvgAList)
  )

df$Age<- ordered(df$Age, levels = c("Young","Middle","Old"))

for(i in unique(df$Age)){
  pick<- df$Age==i
  r<- range(df$x[pick])
  df$x[pick]<- (df$x[pick]-r[1])/(r[2]-r[1])*2-1
}

p <- ggplot(df, aes(x=x,y=AvgDegree, color = Age, group = Age))
p + geom_violin() + geom_jitter() + facet_grid(~ Age)+  geom_boxplot(alpha =0.35, outlier.color = 'NA')+ scale_colour_manual(values = c("red","green","blue"))

```


Two sample t-test for middle and old
```{r}

t.test(middleageAvgAList,oldAvgAList,"greater", var.equal=T)

```


<!-- ###Non-zeros distribution of the 3 age groups: -->

<!-- The sum of number of non-zeros in the average adjacency: -->

<!-- ```{r, echo=FALSE} -->

<!-- df<- data.frame("ROI_Index"= rep(c(1:(n)),3), -->
<!--                 "Age"=as.factor(rep(c("Young","Middle","Old"),each=n)), -->
<!--                 "Number_Non_Zero" = c(rowSums(youngAvgA>0), -->
<!--                                       rowSums(middleageAvgA>0), -->
<!--                                       rowSums(oldAvgA>0)) -->
<!-- ) -->

<!-- df$Age<- ordered(df$Age, levels = c("Young","Middle","Old")) -->

<!-- # df<-df[df$Age!="Young",] -->

<!-- ggplot(df, aes(x=Number_Non_Zero, fill=Age)) + geom_density(alpha = 0.7, bw = 10) -->

<!-- ``` -->

<!-- ```{r,echo=FALSE} -->


<!-- ageGroup<-  (weeks<=63)*1 + (weeks>63 & weeks<=78)*2+(weeks>78)*3 -->
<!-- ageGroup[ageGroup==1] = "Young" -->
<!-- ageGroup[ageGroup==2] = "Middle" -->
<!-- ageGroup[ageGroup==3] = "Old" -->

<!-- df<- data.frame( -->
<!--   "Age"= c("Young","Middle","Old"), -->
<!--   "Number_Non_Zero"=c(sum((youngAvgA>0)*1),sum((middleageAvgA>0)*1),sum((oldAvgA>0)*1)) -->
<!--   ) -->

<!-- df$Age<- ordered(df$Age, levels = c("Young","Middle","Old")) -->


<!-- ggplot(data=df, aes(x=Age, y=Number_Non_Zero, fill=Age)) + -->
<!--     geom_bar(stat="identity") -->

<!-- ``` -->

###Genotypes
###Degree distribution of the 3 genotype groups:

Genotype 0 has less connectivity than group 1 and 2:

```{r,echo=FALSE}

avgAgene0 <- computeAvgA(AdjacencyList[covariates$GENOTYPE==0])
avgAgene1 <- computeAvgA(AdjacencyList[covariates$GENOTYPE==1])
avgAgene2 <- computeAvgA(AdjacencyList[covariates$GENOTYPE==2])

df<- data.frame("ROI_Index_L_R"= rep(c(1:(n)),3),
                "Genotype"=as.factor(rep(c("0","1","2"),each=n)),
                "Degree" = c(rowSums(avgAgene0),rowSums(avgAgene1),rowSums(avgAgene2)),
                "Number_Non_Zero" = c(rowSums(avgAgene0>0),rowSums(avgAgene1>0),rowSums(avgAgene2>0))

)

maxZ<- max(c(avgAgene0,avgAgene1,avgAgene2))

plotHeatmap(avgAgene0, maxZ)+
  ggtitle("Genotype 0")
plotHeatmap(avgAgene1, maxZ)+
  ggtitle("Genotype 1")
plotHeatmap(avgAgene2, maxZ)+
  ggtitle("Genotype 2")


```


The degree distribution:

```{r,echo=FALSE}

ggplot(df, aes(x=Degree, fill=Genotype)) + geom_density(alpha = 0.7, bw = 10)

df2<- data.frame("Genotype"=as.factor((c("0","1","2"))),
                "AvgDegree"=c(sum(avgAgene0),sum(avgAgene1),sum(avgAgene2))/n
)

ggplot(data=df2, aes(x=Genotype, y=AvgDegree, fill=Genotype)) +
    geom_bar(stat="identity")
```

<!-- The non-zeros distribution in the average adjacency: -->

<!-- ```{r,echo=FALSE} -->

<!-- ggplot(df, aes(x=Number_Non_Zero, fill=Genotype)) + geom_density(alpha = 0.7, bw = 10) -->

<!-- df2<- data.frame( -->
<!--   "Genotype"= c("0","1","2"), -->
<!--   "Number_Non_Zero"=c(sum((avgAgene0>0)*1),sum((avgAgene1>0)*1),sum((avgAgene2>0)*1)) -->
<!--   ) -->

<!-- ggplot(data=df2, aes(x=Genotype, y=Number_Non_Zero, fill=Genotype)) + -->
<!--     geom_bar(stat="identity") -->

<!-- ``` -->

###Sex
###Degree distribution of the 2 Sex:

```{r,echo=FALSE}

avgAgender1 <- computeAvgA(AdjacencyList[covariates$GENDER==1])
avgAgender2 <- computeAvgA(AdjacencyList[covariates$GENDER==2])


df<- data.frame("ROI_Index_L_R"= rep(c(1:n),2),
                "Gender"=as.factor(rep(c("1","2"),each=n)),
                "Degree" = c(rowSums(avgAgender1),rowSums(avgAgender2)),
                "Number_Non_Zero" = c(rowSums(avgAgender1>0),rowSums(avgAgender2>0)))

maxZ<- max(c(avgAgender1,avgAgender2))

plotHeatmap(avgAgender1, maxZ)+
  ggtitle("Gender 1")
plotHeatmap(avgAgender2, maxZ)+
  ggtitle("Gender 2")

```

The degree distribution:

```{r,echo=FALSE}

ggplot(df, aes(x=Degree, fill=Gender)) + geom_density(alpha = 0.7, bw = 5)

df2<- data.frame("Gender"=as.factor((c("1","2"))),
                "Avg.Degree"=c(sum((avgAgender1)),sum((avgAgender2)))/n)

ggplot(data=df2, aes(x=Gender, y=Avg.Degree, fill=Gender)) +
    geom_bar(stat="identity")
```

<!-- The non-zeros distribution in the average adjacency: -->

<!-- ```{r,echo=FALSE} -->

<!-- ggplot(df, aes(x=Number_Non_Zero, fill=Gender)) + geom_density(alpha = 0.7, bw = 10) -->

<!-- df2<- data.frame( -->
<!--   "Gender"= c("1","2"), -->
<!--   "Number_Non_Zero"=c(sum((avgAgender1>0)),sum((avgAgender2>0))) -->
<!--   ) -->

<!-- ggplot(data=df2, aes(x=Gender, y=Number_Non_Zero, fill=Gender)) + -->
<!--     geom_bar(stat="identity") -->

<!-- ``` -->


# age x genotype

Genotype 0 (only 4 data points in middle age group, skipped)

```{r}

sum(covariates$GENOTYPE==0)


```


Genotype 1 ( 9 data points across whole age span)

```{r}

sum(covariates$GENOTYPE==1)

pickGene <- covariates$GENOTYPE==1

youngAvgA <- computeAvgA(AdjacencyList[weeks<=63 & pickGene])
middleageAvgA <- computeAvgA(AdjacencyList[weeks>63 & weeks<= 78  & pickGene])
oldAvgA <- computeAvgA(AdjacencyList[weeks>78 & pickGene])

df<- data.frame(
  "Age"= c("Young","Middle","Old"),
  "AvgDegree"=c(sum((youngAvgA)),sum((middleageAvgA)),sum((oldAvgA)))/n
  )

df$Age<- ordered(df$Age, levels = c("Young","Middle","Old"))


ggplot(data=df, aes(x=Age, y=AvgDegree, fill=Age)) +
    geom_bar(stat="identity")


```


```{r,echo=FALSE}

pickGene <- covariates$GENOTYPE==1

youngAvgAList <- getAvgDegree(AdjacencyList[weeks<=63 & pickGene])
middleageAvgAList <- getAvgDegree(AdjacencyList[weeks>63 & weeks<= 78 & pickGene ])
oldAvgAList <- getAvgDegree(AdjacencyList[weeks>78 & pickGene])

n1<- length(youngAvgAList)
n2<- length(middleageAvgAList)
n3<- length(oldAvgAList)


df<- data.frame(
  "x" = runif(n1+n2+n3,-1,1),
  "Age"= as.factor(c(rep("Young",n1),rep("Middle",n2),rep("Old",n3))),
  "AvgDegree"=c(youngAvgAList,middleageAvgAList,oldAvgAList)
  )

df$Age<- ordered(df$Age, levels = c("Young","Middle","Old"))

for(i in unique(df$Age)){
  pick<- df$Age==i
  r<- range(df$x[pick])
  df$x[pick]<- (df$x[pick]-r[1])/(r[2]-r[1])*2-1
}

p <- ggplot(df, aes(x=x,y=AvgDegree, color = Age, group = Age))
p + geom_violin() + geom_jitter() + facet_grid(~ Age)+  geom_boxplot(alpha =0.35, outlier.color = 'NA')+ scale_colour_manual(values = c("green","blue","red" ))
```



Two sample t-test for middle and old
```{r}

t.test(middleageAvgAList,oldAvgAList,"greater", var.equal=T)

```


Genotype 2 ( 12 data points across whole age span)

```{r}

sum(covariates$GENOTYPE==2)

pickGene <- covariates$GENOTYPE==2

youngAvgA <- computeAvgA(AdjacencyList[weeks<=63 & pickGene])
middleageAvgA <- computeAvgA(AdjacencyList[weeks>63 & weeks<= 78  & pickGene])
oldAvgA <- computeAvgA(AdjacencyList[weeks>78 & pickGene])

df<- data.frame(
  "Age"= c("Young","Middle","Old"),
  "AvgDegree"=c(sum((youngAvgA)),sum((middleageAvgA)),sum((oldAvgA)))/n
  )

df$Age<- ordered(df$Age, levels = c("Young","Middle","Old"))


ggplot(data=df, aes(x=Age, y=AvgDegree, fill=Age)) +
    geom_bar(stat="identity")


```


```{r,echo=FALSE}

pickGene <- covariates$GENOTYPE==2

youngAvgAList <- getAvgDegree(AdjacencyList[weeks<=63 & pickGene])
middleageAvgAList <- getAvgDegree(AdjacencyList[weeks>63 & weeks<= 78 & pickGene ])
oldAvgAList <- getAvgDegree(AdjacencyList[weeks>78 & pickGene])

n1<- length(youngAvgAList)
n2<- length(middleageAvgAList)
n3<- length(oldAvgAList)


df<- data.frame(
  "x" = runif(n1+n2+n3,-1,1),
  "Age"= as.factor(c(rep("Young",n1),rep("Middle",n2),rep("Old",n3))),
  "AvgDegree"=c(youngAvgAList,middleageAvgAList,oldAvgAList)
  )

df$Age<- ordered(df$Age, levels = c("Young","Middle","Old"))



for(i in unique(df$Age)){
  pick<- df$Age==i
  r<- range(df$x[pick])
  df$x[pick]<- (df$x[pick]-r[1])/(r[2]-r[1])*2-1
}

p <- ggplot(df, aes(x=x,y=AvgDegree, color = Age, group = Age))

p + geom_violin() + geom_jitter() + facet_grid(~ Age)+  geom_boxplot(alpha =0.35, outlier.color = 'NA')+ scale_colour_manual(values = c("green","blue","red" ))
```



Two sample t-test for middle and old
```{r}

t.test(middleageAvgAList,oldAvgAList,"greater", var.equal=T)

```