```{r,echo=FALSE}
require('igraph')
require('ggplot2')
require('reshape')
require('gridExtra')


setwd("~/git/ABadea/")

listGs<- list.files(path = "./graphml/", pattern = "*.graphml")

#read in covariates and graph list
#find those with common ids, sort by id

covariates<- read.csv("./covariates/predictors.csv",stringsAsFactors = F)
ids <- unlist( lapply(listGs,function(x)strtrim(x,6)))
common_id<- intersect(covariates$RUNNO , ids)

covariates <- covariates[covariates$RUNNO%in%common_id,]
covariates <- covariates[order(covariates$RUNNO),]  

listGs<- listGs[ids%in%common_id]
listGs<- listGs[order(listGs)]

#load and sort the tensor decomp data
load("./tensorDecompBinary.Rda")
load("./listData.Rda")
ids2 <- unlist( lapply(dataList,function(x)strtrim(x$id,6)))
diagC<- tensorDecomp$C[ids2%in% common_id,]
diagC<- diagC[order(ids2[ids2%in% common_id]),]



graphList<- lapply(listGs, function(x){
  read.graph( file = paste("./graphml/",x,sep = ""),format = "graphml")
})

AdjacencyList<- lapply(graphList, function(x){
  get.adjacency(x)
})

HemisphereList<- lapply(graphList, function(x){
  get.vertex.attribute(x,name="hemisphere")
})

```


```{r,echo=FALSE}

colnames(AdjacencyList[[1]])

order_by_hemi<- order(HemisphereList[[1]])


AbyHemiSphere<- (as.matrix(AdjacencyList[[1]]))[order_by_hemi,order_by_hemi]

plotHeatmap<-function(denseA, zlimMax=1){
  rownames(denseA) <- c(1:nrow(denseA))
  colnames(denseA) <- c(1:nrow(denseA))
  m<-melt(denseA)
  n<- nrow(denseA)
  p <- ggplot(m, aes(X1, X2)) +
    geom_tile(aes(fill =  value), colour = "white") + 
    scale_fill_gradient(low = "white",  high = "red",limits=c(0,zlimMax)) +
    theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.y = element_blank()) 
  
  p
}

```

#1. Adjacency matrix view of the 25 subjects

### One random subject:
The adjacency matrix of the whole brain:
the left and right hemispheres look similar with some random difference.

Segments:
Bottom left: *Left Hemisphere*
Upper right: *Right Hemisphere*

```{r,echo=FALSE}
wholeBrain<- plotHeatmap(AbyHemiSphere)
wholeBrain+ geom_hline(yintercept = nrow(AbyHemiSphere)/2 ) + geom_vline(xintercept = nrow(AbyHemiSphere)/2 )
```

### On average:

If we look at them by average, the left and right look almost the same

Segments:
Bottom left: *Left Hemisphere*
Upper right: *Right Hemisphere*


```{r,echo=FALSE}

m<- length(AdjacencyList)


n<- nrow(AdjacencyList[[1]])

computeAvgA<- function(l){
  m<- length(l)
  sumAdj<- matrix(0,n,n)
  for(i in 1:m){
  sumAdj<- sumAdj + (AdjacencyList[[i]])[order_by_hemi,order_by_hemi]
  }
  as.matrix(sumAdj/m)
}

averageAdjacency<- computeAvgA(AdjacencyList)


avgWholeBrain<- plotHeatmap(averageAdjacency)
avgWholeBrain+ geom_hline(yintercept = nrow(AbyHemiSphere)/2 ) + geom_vline(xintercept = nrow(AbyHemiSphere)/2 )
```

```{r,echo=FALSE}

lDegree<- rowSums(averageAdjacency[1:(n/2),1:(n/2)])
rDegree<- rowSums(averageAdjacency[(n/2+1):(n),(n/2+1):(n)])



df<- data.frame("ROI_Index"= rep(c(1:(n/2)),2),
                "Hemisphere"=as.factor(rep(c("L","R"),each=n/2)),
                "Degree" = c(lDegree,rDegree)
)

```

### Compare left and right hemispheres in average

The degree plots conditional on the L and R hemisphere also confirms that the left and right are similar:

Degree plot by ROI index
```{r,echo=FALSE}
ggplot(data=df,
       aes(x=ROI_Index, y=Degree, colour=Hemisphere)) +
  geom_line()
```

Degree distribution
```{r,echo=FALSE}
ggplot(df, aes(x=Degree, fill=Hemisphere)) + geom_density(alpha = 0.7, bw = 3)
```


#2. Age related degeneration


### The degree distribution over age (weeks)

```{r}

weeks<- covariates$AGE_WEEKS

degreeForAll<-c(sapply(AdjacencyList,function(x){
  rowSums(as.matrix(x))
}))


df<- data.frame("ROI_Index_L_and_R"= rep(c(1:n),m),
                "Age"= as.factor(rep(weeks,each=n)),
                "Degree" = degreeForAll)


ggplot(data=df,
       aes(x=ROI_Index_L_and_R, y=Degree, colour=Age)) +
  geom_line()

ggplot(df, aes(x=Degree, fill=Age)) + geom_density(alpha = 0.7, bw = 5)


```



```{r}


sumAForAll<-c(sapply(AdjacencyList,function(x){
  sum(as.matrix(x))
}))

sumNNZForAll<-c(sapply(AdjacencyList,function(x){
  sum(as.matrix(x>0))
}))


degreeForAll<-(lapply(AdjacencyList,function(x){
  rowSums(as.matrix(x))
}))

avgDegreeNZ<- sapply(degreeForAll, function(x){mean(x[x>0])})


df<- data.frame("Age"= weeks,
                "SumA" = sumAForAll,
                "AvgDegree" = avgDegreeNZ,
                "GenoType" = as.factor(covariates$GENOTYPE),
                "Gender" = as.factor(covariates$GENDER)
                )

ggplot(data=df,
       aes(x=Age, y=SumA,col=GenoType)) +
  geom_line()+ggtitle("Number of Edges vs Age")

ggplot(data=df,
       aes(x=Age, y=SumA,col=Gender)) +
  geom_line()+ggtitle("Number of Edges vs Age")




ggplot(data=df,
       aes(x=Age, y=avgDegreeNZ,col=GenoType)) +
  geom_line()+ggtitle("Average Degree vs Age")



ggplot(data=df,
       aes(x=Age, y=avgDegreeNZ,col=Gender)) +
  geom_line()+ggtitle("Average Degree vs Age")



```

### Divide age into 3 groups

We use 63 and 78.0 as cutoffs to divide mice into three groups: *young*,*middle* and *old*.
```{r}
plot(weeks)
abline(h=63,lty=2)
abline(h=78,lty=2)

quantile(weeks,c(0.25,0.75))
```

### Comparing Adjacency matrices

Obviously, *old group seems significantly more degenerate than the young and middle*, while the difference beween young and middle is more subtle.

```{r,echo=FALSE}

youngAvgA <- computeAvgA(AdjacencyList[weeks<=63])
middleageAvgA <- computeAvgA(AdjacencyList[weeks>63 & weeks<= 78  ])
oldAvgA <- computeAvgA(AdjacencyList[weeks>78])

zlim<- max(c(youngAvgA, middleageAvgA, oldAvgA))

youngAvgWholeBrain<- plotHeatmap(youngAvgA,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("Young")

middleageAvgWholeBrain<- plotHeatmap(middleageAvgA,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("Middle")

oldAvgWholeBrain<- plotHeatmap(oldAvgA,zlimMax = zlim) + geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("old") 

```


Young:
```{r,echo=FALSE}
sum(weeks<=63)
youngAvgWholeBrain
```

Middle:
```{r,echo=FALSE}
sum(weeks>63 & weeks<= 78)

middleageAvgWholeBrain

```

Old:
```{r,echo=FALSE}
sum(weeks> 78)
oldAvgWholeBrain
```

The distance between Middle and Young:

```{r}
sqrt(sum((middleageAvgA-youngAvgA)^2))
```


The distance between Old and Middle:

```{r}
sqrt(sum((oldAvgA-middleageAvgA)^2))
```


###Dichotomizing the average plots shows a clearer difference

The dichotomized view with connecting probability>0: it shows the old clearly has less connectivity than the middleage and young.

```{r,echo=FALSE}

zlim<- 1

youngAvgWholeBrainDicho<- plotHeatmap((youngAvgA>0)*1,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("Young")

middleageAvgWholeBrainDicho<- plotHeatmap( (middleageAvgA>0)*1,zlimMax = zlim)+ geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("Middle")

oldAvgWholeBrainDicho<- plotHeatmap((oldAvgA>0)*1,zlimMax = zlim) + geom_hline(yintercept = n/2 ) + geom_vline(xintercept =  n/2 )+
  ggtitle("old") 

youngAvgWholeBrainDicho
middleageAvgWholeBrainDicho
oldAvgWholeBrainDicho

```

The count of the vectices with probability > 0 in the average plot

```{r,echo=FALSE}


ageGroup<-  (weeks<=63)*1 + (weeks>63 & weeks<=78)*2+(weeks>78)*3
ageGroup[ageGroup==1] = "Young"
ageGroup[ageGroup==2] = "Middle"
ageGroup[ageGroup==3] = "Old"

df<- data.frame(
  "Age"= c("Young","Middle","Old"),
  "Count"=c(sum((youngAvgA>0)*1),sum((middleageAvgA>0)*1),sum((oldAvgA>0)*1))
  )

ggplot(data=df, aes(x=Age, y=Count, fill=Age)) +
    geom_bar(stat="identity")
    
```

###Degree distribution of the 3 age groups:

Middle and young mice have higher level of connectivy  (degree) than the old.


```{r,echo=FALSE}

df<- data.frame("ROI_Index"= rep(c(1:(n)),3),
                "Age"=as.factor(rep(c("Young","Middle","Old"),each=n)),
                "Degree" = c(rowSums(youngAvgA),rowSums(middleageAvgA),rowSums(oldAvgA))
)

df<-df[df$Age!="Young",]

ggplot(df, aes(x=Degree, fill=Age)) + geom_density(alpha = 0.7, bw = 6)

```


```{r, echo=FALSE}


df<- data.frame("ROI_Index"= rep(c(1:(n)),3),
                "Age"=as.factor(rep(c("Young","Middle","Old"),each=n)),
                "Degree" = c(rowSums(youngAvgA>0),rowSums(middleageAvgA>0),rowSums(oldAvgA>0))
)

df<-df[df$Age!="Young",]

ggplot(df, aes(x=Degree, fill=Age)) + geom_density(alpha = 0.7, bw = 20)

```


```{r}
sum(youngAvgA>0)
sum(middleageAvgA>0)
sum(oldAvgA>0)

mean(youngAvgA[youngAvgA>0])
mean(middleageAvgA[middleageAvgA>0])
mean(oldAvgA[oldAvgA>0])

```

#Use of dimension reduction as an more wieldly alternative to adjaceny matrices:

We applied joint embedding on the 25 subjects and extract the diagonal distinct vectors

### Pairs plot of the first 5 dimensions:
```{r,echo=FALSE}

require('GGally')


df<- data.frame("Latent Dimension Idx"= rep(c(1:30),m),
                "Age"=as.factor(rep(ageGroup,each=30)),
                "Latent Coordinate" = c(t(diagC))
)



K<- 5

seleDf<- diagC[,1:K]


# seleDf<- cbind((seleDf))

colnames(seleDf)<- c(  sapply(c(1:K),function(x){paste("x",x,sep = ".")}))

newDf<- as.data.frame(seleDf)
newDf<- cbind(newDf, ageGroup)
ggpairs(newDf,aes(col=ageGroup,alpha=0.4, binwidth=1))


```


###Average scree plots of the 3 groups:
From the low dimension representation, *old group looks distinct from the middle and young groups*. This is consistent with the finding in the adjacency plots. This suggests that one can use the low dimensional representation as a surrogate for the large adjacency matrix.

```{r,echo=FALSE}

C1<- diagC[ageGroup=="Young",]
C2<- diagC[ageGroup=="Middle",]
C3<- diagC[ageGroup=="Old",]

df<- data.frame("Latent.Dim"= rep(c(1:(30)),3),
                "AgeGroup"=as.factor(rep(c("Young","Middle","Old"),each=30)),
                "Latent.Coor" = c(colMeans(C1),colMeans(C2),colMeans(C3))
)

ggplot(data=df,
       aes(x=Latent.Dim, y=Latent.Coor, colour=AgeGroup)) +
  geom_line()

```




###Genotypes
###Degree distribution of the 3 genotype groups:

Genotype 0 has less connectivity than group 1 and 2:

```{r,echo=FALSE}

avgAgene0 <- computeAvgA(AdjacencyList[covariates$GENOTYPE==0])
avgAgene1 <- computeAvgA(AdjacencyList[covariates$GENOTYPE==1])
avgAgene2 <- computeAvgA(AdjacencyList[covariates$GENOTYPE==2])


df<- data.frame("ROI_Index_L_R"= rep(c(1:(n)),3),
                "Genotype"=as.factor(rep(c("0","1","2"),each=n)),
                "Degree" = c(rowSums(avgAgene0),rowSums(avgAgene1),rowSums(avgAgene2))
)

maxZ<- max(c(avgAgene0,avgAgene1,avgAgene2))

plotHeatmap(avgAgene0, maxZ)+
  ggtitle("Genotype 0")
plotHeatmap(avgAgene1, maxZ)+
  ggtitle("Genotype 1")
plotHeatmap(avgAgene2, maxZ)+
  ggtitle("Genotype 2")


sum(avgAgene0>0)
sum(avgAgene1>0)
sum(avgAgene2>0)

mean(avgAgene0[avgAgene0>0])
mean(avgAgene1[avgAgene1>0])
mean(avgAgene2[avgAgene2>0])


ggplot(df, aes(x=Degree, fill=Genotype)) + geom_density(alpha = 0.7, bw = 1)

```



###Sex
###Degree distribution of the 2 Sex:



```{r,echo=FALSE}

avgAgender1 <- computeAvgA(AdjacencyList[covariates$GENDER==1])
avgAgender2 <- computeAvgA(AdjacencyList[covariates$GENDER==2])


df<- data.frame("ROI_Index_L_R"= rep(c(1:n),2),
                "Gender"=as.factor(rep(c("1","2"),each=n)),
                "Degree" = c(rowSums(avgAgender1),rowSums(avgAgender2))
)

maxZ<- max(c(avgAgender1,avgAgender2))

plotHeatmap(avgAgender1, maxZ)+
  ggtitle("Gender 1")
plotHeatmap(avgAgender2, maxZ)+
  ggtitle("Gender 2")


ggplot(df, aes(x=Degree, fill=Gender)) + geom_density(alpha = 0.7, bw = 2)

```


